%!TEX root = thesis.tex

\chapter{Conclusions \& Ignition Model}
\label{chap:conclusion}
The preceding chapters of this document presented findings from studies evaluating the influence of various parameters on the probability of ignition of fuel beds in wild fires. The specific objectives of this work are as follows:
        \begin{enumerate}
            \item Determine the effects of particle morphology on ignition propensity
            \item Ascertain ignition dependence on heating location(s), mode, and rate
            \item Identify and quantify the influence of environmental conditions on ignition propensity.
            \item Identify the influence of fuel bed chemical composition on flaming ignition.
        \end{enumerate}
The conclusions of each effort are restated in the following sections followed by a discussion of the implications of the results as they pertain to each specific objective and the overall objective of this work.

\section{Sensitivities of Porous Beds and Plates to Ignition by Firebrands}
        \begin{enumerate}
            \item 
                Smaller particles ignite more readily in porous beds than larger particles when heat transfer from the heater is primarily through conduction. This was evident by higher ignition probabilities, in general, of the smaller particles for a fixed heater temperature. As particle sizes increase radiant heat transfer becomes more important and fuel beds with larger particles were more likely than smaller particles to ignite at extended times (\textgreater 100\si{\second}) due to the increased importance of radiant ignition. 
            \item
                Douglas-fir plates ignite at times where conduction is the dominant mode of heat transfer (\textless 10\si{\second}) due to the higher thermal conductivity of the solid plates. The ignition probability of plates was the most similar to the larger particle, in particular at lower heater temperatures, due to dispersed heating of the porous fuel bed through radiation and the increased thermal conductivity of the plates creating similar temperature profiles. The rise in ignition probability  over a smaller heater temperature range time with temperature results from more consistent contact between the heater and plate surface.
            \item 
                Heat flux delivered to the fuel bed, when compared to heater temperature, is more indicative of ignition likelihood and ignition time for porous fuel beds. Heat flux is a more significant predictor of ignition because it captures differences in heat transfer modes and particle contact that heater temperature values do not. While this finding is not new, what is novel is that the mixed mode of heating (conduction and radiation) has a significant impact on the flaming ignition of fuel beds.
            \item 
                Consideration of the transport characteristics of pyrolyzate gases near the high temperature source can be important for more fully predicting ignition propensity. A \textit{Da} of ignition, in relation to the measured heat flux and thermal diffusivity of the fuel beds, is a promising relationship for predicting ignition for the porous fuel beds.  
        \end{enumerate}
    Further work is needed to verify that the \textit{Da} may be used to predict ignition for solid surfaces and for porous fuel beds with varying chemical compositions.  If proven valid, the (\textit{Da}), measured/predicted heat fluxes, and fuel bed properties may be used to help predict ignition of fuel beds both in and out of the WUI, ultimately helping to increase the effectiveness of fire prevention and suppression efforts.

\section{Influence of Wind on Flaming Ignition of Porous Wood Fuel Beds}
        \begin{enumerate}
            \item
            An increase in wind speed above quiescent conditions reduces the temperature required for the flaming ignition of a fuel bed. For example, an increase in wind speed of 3.5\si{\meter\per\second} from quiescent increases the ignition probability of a fuel bed from under 30\% to roughly 100\%. However, a linear increase in wind speed does not result in a linear increase in ignition probability. Thresholds in wind speed exist above which temperatures required to achieve ignition actually increase. For example, when the wind is oriented 45\si{\degree} from the heater centerline increasing the wind from quiescent to 3.5\si{\meter\per\second} reduces the temperature required for ignition probability by 30\%. However, increasing the wind speed from quiescent to 5.8\si{\meter\per\second} reduces the temperature required for ignition by only 25\%. Presumably these thresholds occur because of reductions in residence time. 
            
            \item The temperatures at which ignition occurs for porous fuel beds is sensitive to the orientation of a firebrand relative to the wind direction. Higher temperatures are typically required for ignition for a heater parallel to the flow compared to 45\si{\degree} and perpendicular to the flow. This sensitivity attributed to differences in recirculation zones and residence times of air and pyrolyzates near the hottest region of the heater. Thus, predictions of ignition probabilities that consider wind may need to include both wind speed and orientation to obtain sufficient accuracy.
            
            \item Times to flaming ignition of porous fuel beds are sensitive to the firebrand/heater angle in the presence of wind.
            The parallel heater orientation ignites at the longest time followed by the 45\si{\degree} case with the perpendicular cases igniting in the shortest amount of time. High speed images indicate that ignition typically occurs in regions where recirculation zones occur, as shown in CFD calculations. The heightened propensity to ignition is attributed to increased residence times of pyrolyzates in the recirculation zones as supported by calculations.
        \end{enumerate}
    The conclusions of this work show that ignition is favored when a firebrand(s) land on a fuel bed under wind speeds and orientations that promote greater residence times of pyrolyzates near a high temperature region of firebrands. It was observed that increases of wind speed, of a magnitude that may commonly occur during wildfires, can increase the probability of fuel bed ignition from very unlikely to a near certainty regardless of the ember orientation to the wind. This highlights the increased risk of spot fires due to ignition of fuel beds that accompanies wind in a wildfire. 

\section{Effect of Fuel Bed Composition on Flaming Ignition Probability}
        \begin{enumerate}
            \item Of the six materials tested, only three were able to be ignited within below the 750\si{\celsius} maximum temperature of the firebrand surrogate. Two of the three materials for which ignition was not observed were bark. The higher ignition threshold of the bark materials is attributed to higher lignin concentrations. The resistance to ignition present in these materials suggest that using materials high in lignin in locations at risk for firebrand attack in wildfires may reduce ignitions and subsequent losses of structures. It is acknowledged that these results are for flaming ignition only, and high lignin materials may pose a significant risk if smoldering or the smoldering to flaming transition occurs.
            
            \item For Douglas-fir and pine wood, an increase in wind speed decreased the temperature required for ignition. However, for oak wood, an increase in wind speed inhibited ignition. Increases in ignition propensity with an increase in wind are attributed to the formation of recirculation zones near the ignition source. In the case of oak wood, the formation of recirculation zones inhibits ignition. The root cause of the dichotomy in ignition trends is unclear; however, differences in the flammability of pyrolysis products corresponding to changes in chemical composition are most likely.
            
            \item A direct correlation between the fuel bed composition and the ignition propensity was not observed. This is attributed to the complexity of biomass pyrolysis. However, a local minimum ignition threshold near the composition of Douglas-fir wood is present that warrants further investigation. 
        \end{enumerate}
    The conclusions of this work show that ignition is sensitive to the chemical composition of fuel beds in both quiescent conditions and under wind. An increase in wind was observed to increase ignition probability for some materials; however, this was not universally the case. Materials high in lignin were among the least likely to ignite and more suitable for placement in areas where firebrands are likely to land or accumulate in a WUI environment. Further and more targeted analysis of the differences in pyrolysis products for each material will likely yield further insight. 

\section{Influence of Multiple Firebrands on the Ignition of Fuel Beds}
        \begin{enumerate}
            \item Increasing the wind speed lowers the threshold for ignition. However, the magnitude of the change in ignition threshold is sensitive to the spacing and temperatures of cartridge heaters used. The ignition threshold may reduce between 20\% and 60\% depending on the heater configurations.
            
            \item At wind speeds of 5.8\si{\meter\per\second} the ignition threshold is largely independent of thermal interactions between firebrand surrogates. The lack of sensitivity to inter-firebrand effects is attributed to the ignition being largely controlled by the fluid dynamics around the firebrand surrogates. Ignition is largely controlled by the propensity of pyrolysis products to accumulate and is independent of thermal interactions with nearby objects, whether they are energy sources or sinks. 
            
            \item In low wind conditions (e.g., 0.5\si{\meter\per\second}), the ignition threshold is more sensitive to thermal interactions between the firebrand surrogates than in higher wind conditions. For example, when the firebrand was spaced one diameter apart, the difference between ignition thresholds when the upstream firebrand is inert or an energy source is 34\%. When the firebrand surrogates were spaced such that they did not thermally interact (five diameters apart), the difference in ignition threshold was 4\%. Thus, at low wind speeds or in quiescent conditions, ignition is sensitive to thermal interactions with nearby objects. Ignition is promoted if nearby objects supply energy and inhibited if thermal sinks are nearby. 
            
            \item In configurations where the firebrand is downstream of a flow obstacle, ignition may occur via the smoldering to flaming transition. The smoldering to flaming transition appears to be facilitated by either accumulation of pyrolysis gases in the downstream edge of the flow obstruction or by the formation of an overhang as the fuel bed recedes underneath the obstruction. This phenomenon produces ignition at similar temperatures to other configurations and may not be important for ignition predictions but may be of interest to smoldering research. 
        \end{enumerate}
    The conclusions of this work show that ignition of a fuel bed is more likely when a firebrand or firebrand lands on a fuel bed under conditions that promote recirculation zones near locations where pyrolysis gases are released from the fuel bed. It was observed that under windy conditions ignition is less sensitive to nearby embers or inert objects than at low wind speeds. These observed sensitivities highlight the increased risk of fuel bed ignition in windy conditions and in cases where firebrands may accumulate in all wind conditions. 

\section{Ignition Model}
    The overall objective of this work is to identify parameters and processes that control the ignition of a fuel bed when an ember lands on it with an anticipated impact of enabling the creation of a simplified model that may be used to determine the likelihood of ignition quickly. A model was created to predict ignition to evaluate the effectiveness of this work in meeting those goals. The model was then applied to the results of other studies that utilized different fuel bed materials and ignition sources to determine cross-study applicability.
    
    After the completion of the experimental efforts outlined in the previous chapters the results were aggregated into a single dataset containing the results from the 1086 tests conducted. The parameters included in the model and the corresponding units are as follows:
        \begin{itemize}
            \item Wind speed (\si{\meter\per\second)}
            \item Wind direction relative to longitudinal firebrand axis (\si{\degree})
            \item Ratio of characteristic lengths between particle size and ember size (-)
            \item Chemical composition: mass fractions of Cellulose, Hemicellulose, Lignin, Tannins, and Triglycerides (-)
            \item Fuel bed bulk density (\si{\kilo\gram\per\cubic\meter})
            \item Fuel bed thermal diffusivity (\si{\square\meter\per\second})
            \item Average temperature of the firebrand(s) (\si{\celsius})
            \item Number of firebrands (-)
            \item Spacing between firebrands as multiple of characteristic length (-)
        \end{itemize}
    The aggregated data was then used to create a Random Forest Classifier model using the scikit-learn package in python~\cite{scikit-learn} to estimate ignition outcomes for each test. A 0.75/0.25 test-train split was used for training the models, meaning that 25\% of the data was used to train the model and the remaining 75\% was used to test the model accuracy. The R\textsuperscript{2} validation score for the model with an out-of-bag score of 0.83, which means that the model predicts ignition correctly 83\% of the time. While the accuracy of the model is likely acceptable if such a model was to be used as an indicator or predictor of ignition during a fire or to inform preventative measures, it would be beneficial to reduce the number of inputs required. An examination of the most important input parameters and refinement of the model found that accuracy was not degraded if the number of parameters was reduced to include only the following variables:
        \begin{itemize}
            \item Average temperature of the firebrands
            \item Fuel bed bulk density
            \item Wind speed
            \item Ratio of characteristic lengths between particle size and ember size
            \item Wind direction
        \end{itemize}
    Reducing the model to include only the five characteristics above maintained the 83\% prediction accuracy. While it is beneficial to have an accurate predictive model for a single set of experiments in a similar apparatus, a model capable of predictions across different is far more valuable for preventing home loss. The model was applied to select studies from the literature to evaluate the accuracy across studies. The authors and relevant parameters from each study are shown in Table~\ref{tab:otherStudies}. Three of the studies presented in Table~\ref{tab:otherStudies} used hot metal particles dropped onto fine fuel beds of processed material. The studies by Hadden et al. and Urban et al. used pure cellulose as the fuel bed, and the study by Zak et al. used a blend of grasses processed into a powder. The fourth study, conducted by Filkov et al., used multiple bark embers of various sizes dropped onto dried, but otherwise unprocessed, pine needle duff. A wind direction of 0\si{\degree} was used for for the not metal particle studies since the geometry of the metal particles is anticipated to produce recirculation zones similar to that of a cylindrical heater parallel to the flow. A wind direction if 45\si{\degree} was used for the pine bark embers since the embers provided a more significant flow disruption than the metal particles but were dropped randomly and therefore not considered to be normal to the flow in all cases. The average temperature was defined as the average of the initial temperature and ambient temperature for the hot metal particles. More sophisticated estimates of the average temperature based on the average time to ignition and estimated heat losses from the ember were conducted, but the model accuracy was not improved. Thus, a simpler metric was chosen. The average temperature of the bark embers was estimated from measurements of similar embers in the literature~\cite{Fateev2017a, Matvienko2018}.
        \begin{table}[hbpt]
            \centering
            \caption{Data inputs for the model from selected studies.}
            \begin{tabular}{crrrrrc}
                Authors & $\bar{T}$ (\si{\celsius}) &  $\bar{\rho}_{bed}$ (\si{\kilo\gram\per\cubic\meter})& $U$ (\si{\meter\per\second}) & $L_{p}/L_{e}$ (-)& $\Theta_{wind}$ (\si{\degree}) & Citation\\
                \hline
                 Hadden et al. & 260 - 560 & 200 & 0.5       & 0.02 - 0.45 & 0  & \cite{Hadden2011}\\
                 Urban et al.  & 285 - 560 & 282 & 0.5       & 0.06 - 0.25 & 0  & \cite{Urban2018}\\
                 Zak et al.    & 310 - 560 & 338 & 0.5       & 0.04 - 0.15 & 0  & \cite{Zak2014} \\
                 Filkov et al. & 407       & 105 & 1, 1.5, 2 & 0.17 - 2.60 & 45 & \cite{Filkov2016}
            \end{tabular}
            \label{tab:otherStudies}
        \end{table}
    The first approach used was to apply a reduced model created from only the data generated from studies in this dissertation and predict ignition for each of the studies in the literature. The accuracy of the model at predicting ignition for other studies was 50\%, which is entirely unusable for any type of predictive modeling. However, when the model was retrained using the same five parameters but including 25\% subsets of the Zak et al. and the Filkov et al. data, in conjunction with the results of this work, the predictive accuracy increased to 98\% for the Zak et al. study, 94\% for the Hadden et al. study, 81\% for the Urban et al. study and 74\% for the Filkov et al. study. The increase in accuracy suggests that predictions can be made across differences in experimental conditions, ember types, and fuel bed types using as few as five parameters. Predictions across experimental conditions have been one of the biggest hurdles to creating a general predictive model. Identifying the primary controlling parameters of ignition in this work has created a framework for bridging the predictive gap across various configurations.  