% !BIB TS-program = biber
\documentclass[onehalf,11pt]{beavtex/beavtex}

\input{thesis-header.tex}

% things to check or replace
% replace your name here.
\newcommand{\TheAuthors}{Derek Bean}
\title{Ignition of Fuel Beds by Firebrands}
\degree{Doctor of Philosophy}
\major{Mechanical Engineering}

\submitdate{June 7, 2022}
\commencementyear{2022}

\author{\TheAuthors{}}
\doctype{Thesis}
\department{Mechanical, Industrial, and Manufacturing Engineering}
\depttype{School}
\depthead{Head}

\advisor{David Blunck}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% figure location
\graphicspath{{./Figures/}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract
{
The severity of wildfires around the globe is increasing. At the same time, urban development is expanding outward into areas where severe fires occur. There is an increased risk of home loss to fires in areas where severe fires and urban expansion meet. Ignition of homes or nearby fuel is a significant mechanism of home loss. To date, a general model for ignition has proven elusive due to the limited quantification of parameters that control ignition. Thus, understanding and quantifying the processes and parameters that control ignition is essential to reducing home losses. This work considered the influence of fuel bed properties (i.e., particle size and chemical composition), environmental conditions near the ignition site (i.e., wind speed and direction), and variations of ignition sources (i.e., spacing, count, and energy deposition) on the likelihood of ignition. A combination of experimental and computational approaches were conducted to determine the influence of the parameters studied on ignition. Ignition propensity increased when conduction was favored due to small particle sizes compared to radiation-driven heat transfer to fuel beds with larger particle sizes. Ignition was also sensitive to fuel bed species. Fuels high in lignin were not able to be ignited, and increased wind speed decreased the ignition threshold for some, but not all, materials. The formation of recirculation zones caused by wind decreased the ignition threshold. The propensity of wind orientation with respect to the ignition source influenced the formation of recirculation zones. As the size of the recirculation zone increased, the ignition threshold decreased. In windy conditions, the presence of multiple firebrands had little influence on the ignition threshold. However, at low wind speeds, interactions between additional firebrands significantly influenced ignition. Finally, results from all ignition tests conducted in this work were aggregated, and an ignition model was created. It is anticipated that the insights gained from these studies and the subsequent model may act as a novel framework for predicting ignition. 
}


\acknowledgements{
I have been fortunate to be perpetually surrounded by people who have given me an immense amount of support and guidance throughout my life. This has been especially true in my studies. This period of my life has been truly transformative. It would not have been possible without all of the wonderful people in my life. I owe the greatest thanks to my family, whose unwavering support, encouragement, and guidance have been invaluable. I would not be who I am or where I am today if it was not for you all. It is impossible to express how much I appreciate everything you have done for me. To Heather, thank you for being my partner through the toughest parts of this journey, for asking so many great questions, and for encouraging me to keep going so that I may reach my goals. I doubt I would have made it this far without you. 

Thank you to Dr. David Blunck for giving me the opportunity to conduct this work. With the graduate school experience you have facilitated, I have achieved things I did not know I was capable of. I have the deepest gratitude for your guidance, insight, and the example that you set. I owe a similar debt of gratitude to my friends and colleagues. I have made many lifelong friendships and created many fond memories. Thank you all for guiding me to who I am today, and I am excited to see where life takes us.

This work was supported by the National Institute of Standards and Technology as part of project number 70NANB17H281.
}


%\dedication{(optional dedication)}

\begin{document}

\maketitle

\mainmatter

% By including `''%!TEX root = thesis.tex` in the top of each included file,
% most LaTeX editors will automatically compile the main package when you are working
% within a file.
%\printnomenclature
\include{introduction}
\include{litreview}
\include{manuscript1}
\include{manuscript2}
\include{manuscript3}
\include{manuscript4}
%\include{results}
\include{conclusions}

\addcontentsline{toc}{chapter}{Bibliography}

\newpage

\printbibliography[]


%% INCLUDE IF YOU NEED AN APPENDIX

% \newpage
%\appendix
%\input{appendixA.tex}

\end{document}
